---
title: "Check if Bovy's Matrices Actually Do Only Span Same Space"
author: "David Gerard"
date: "March 21, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Abstract

I want to know if Bovy's EM algorithm will only result in estimate that span the same space as the initial conditions. Turns out this is correct. We just get a scaled estimate of the initial vector in the rank 1 case.

# Analysis

First, I generate some fake data

```{r}
set.seed(25)
n <- 100
p <- 10
ydata <- matrix(rnorm(n * p), nrow = n)
ycovar <- list()
ycov_list <- lapply(rep(p, n), FUN = stats::rchisq, df = 1)
for (index in 1:n) {
  ycovar[[index]] <- diag(ycov_list[[index]])
}

K <- 3 ## number of mixtures
xamp <- runif(K) ## initial mixture proportions
xamp <- xamp / sum(xamp)

## Here, I assume that observed and true data have same dimension
xmean <- matrix(0, nrow = K, ncol = p) ## initial mixture means

xcovar <- list() ## initial covariance matrices

v1 <- matrix(rnorm(p), ncol = 1)
xcovar[[1]] <- tcrossprod(v1) ## the rank 1 psd matrix

xcovar[[2]] <- diag(p) ## full rank identity

V3 <- matrix(rnorm(p * 5), ncol = 5) ## rank 5 matrix
xcovar[[3]] <- tcrossprod(V3)
```

Now run extreme deconvolution. 

```{r, error=TRUE}
library(ExtremeDeconvolution)

eout <- extreme_deconvolution(ydata = ydata, ycovar = ycovar,
                              xamp = xamp, xmean = xmean, xcovar = xcovar,
                              fixmean = TRUE)

```

Check to see if estimate is in same span as initial value.

```{r}
eigen_out <- eigen(eout$xcovar[[1]])
v1hat <- eigen_out$vectors[, 1]

v1hat / v1
```

Yes, it is. v1hat is just a scalar multiple of v1.

What about the rank 5 matrix. If they are in the same span, then the last five columns should all be zero (since rank 5 and this is they form a basis for the nullspace).

```{r}
qrout <- qr(eout$xcovar[[3]])
Q <- qr.Q(qrout, complete = TRUE)

t(V3) %*% Q
abs(t(V3) %*% Q) < 10 ^ -8
```

```{r}
sessionInfo()
```
