---
title: "Check if UltimateDeconvolution's Matrices Also Span Same Space"
author: "David Gerard"
date: "March 22, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Abstract

Recall from [here](check_bovy.html) that Extreme Deconvolution only returns estimates in the same spann as the initial values. I want to check if my new [UltimateDeconvolution](https://github.com/dcgerard/UltimateDeconvolution) also has this property. It turns out that it does not --- which is good! I also show that the likelihood increases at every fixed point iteration.

# Generate Data and Run Ultimate Deconvolution

First, I generate some fake data

```{r}
set.seed(25)
n <- 100
p <- 10
x_mat <- matrix(rnorm(n * p), nrow = n)
s_mat <- matrix(stats::rchisq(n * p, df = 1), nrow = n)

K <- 3 ## number of mixtures
pi_vec<- runif(K) ## initial mixture proportions
pi_vec <- pi_vec / sum(pi_vec)
v_mat <- matrix(rnorm(p*K), ncol = K)

```

Now run ultimate deconvolution. 

```{r, cache=TRUE}
library(UltimateDeconvolution)

udout <- ultimate_deconvolution(x_mat = x_mat,
                                s_mat = s_mat,
                                v_mat = v_mat,
                                pi_vec = pi_vec)

```

# Check that Likelihood Increases Every Iteration

First, just check that the likelihood increases every iteration. I know it does because I used an assertion for this, but it will be good to plot this so that I remember.

```{r}
suppressMessages(library(tidyverse))
dat <- data_frame(Iteration = 1:length(udout$llike_vec),
                  Loglikelihood = udout$llike_vec)
ggplot(data = dat, mapping = aes(x = Iteration, y = Loglikelihood)) +
  geom_line() +
  theme_bw()
```

# Not in Same Spann as Initial Value.

Check to see if estimate is in same span as initial value. It is not (if it were then each column would contain repetitions of the same numers).

```{r}
v_mat / udout$v_mat
```


```{r}
sessionInfo()
```
