---
title: "Check likelihoods"
author: "David Gerard"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Abstract

Here, I look at whether ExtremeDeconvolution or UltimateDeconvolution returns higher likelihood values. I am strangely getting that ExtremeDeconvolution is making the likelihood *decrease*. Am I coding it up wrong?

# Generate Data and Run Ultimate Deconvolution

First, I generate some fake data

```{r}
set.seed(25)
n <- 11
p <- 13
x_mat <- matrix(rnorm(n * p), nrow = n)
s_mat <- matrix(stats::rchisq(n * p, df = 1), nrow = n)

K <- 3 ## number of mixtures
pi_vec<- runif(K) ## initial mixture proportions
pi_vec <- pi_vec / sum(pi_vec)
v_mat <- matrix(rnorm(p*K), ncol = K)
```

And I reformat it for ExtremeDeconvolution

```{r}
s_list <- list()
for (index in 1:n) {
  s_list[[index]] <- diag(s_mat[index, ])
}

U_list <- list()
for (index in 1:K) {
  U_list[[index]] <- tcrossprod(v_mat[, index])
}
```

Now run `ultimate_deconvolution`.

```{r}
library(UltimateDeconvolution)

udout <- ultimate_deconvolution(x_mat = x_mat,
                                s_mat = s_mat,
                                v_mat = v_mat,
                                pi_vec = pi_vec)

```

And I run `extreme_deconvolution`.

```{r}
library(ExtremeDeconvolution)
edout <- extreme_deconvolution(ydata = x_mat,
                               ycovar = s_list,
                               xamp = pi_vec,
                               xmean = matrix(0, nrow = K, ncol = p),
                               xcovar = U_list,
                               fixcovar = FALSE,
                               fixamp = FALSE)
```

Get out estimates from both fits.

```{r}
ed_pivec <- edout$xamp
ed_vmat <- matrix(NA, nrow = p, ncol = K)
for (index in 1:K) {
  eout <- eigen(edout$xcovar[[index]])
  ed_vmat[, index] <- eout$vectors[, 1] * sqrt(eout$values[1])
  
  assertthat::assert_that(all(abs(tcrossprod(ed_vmat[, index]) - edout$xcovar[[index]]) < 10 ^ -12))
}

ud_pivec <- udout$pi_vec
ud_vmat  <- udout$v_mat
```

Again, ExtremeDeconvolution just returns a scalar multiple of the start values

```{r}
ed_vmat / v_mat
```

While UltimateDeconvolution does not

```{r}
ud_vmat / v_mat
```

Now we compare log-likelihoods

```{r}
initial_llike <- UltimateDeconvolution:::dmixlike(x_mat = x_mat,
                                                  s_mat = s_mat,
                                                  v_mat = v_mat,
                                                  pi_vec = pi_vec,
                                                  log = TRUE)

ed_llike <- UltimateDeconvolution:::dmixlike(x_mat = x_mat,
                                             s_mat = s_mat,
                                             v_mat = ed_vmat,
                                             pi_vec = ed_pivec,
                                             log = TRUE)

ud_llike <- UltimateDeconvolution:::dmixlike(x_mat = x_mat,
                                             s_mat = s_mat,
                                             v_mat = ud_vmat,
                                             pi_vec = ud_pivec,
                                             log = TRUE)

initial_llike

ed_llike

ud_llike
```

The likelihood from UltimateDeconvolution is `r exp(ud_llike - ed_llike)` times greater than the likelihood from ExtremeDeconvolution.

```{r}
sessionInfo()
```
